---
title: "hw-05"
output: github_document
 
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(gapminder)
library(ggplot2)
library(plotly)
library(tidyverse)
library(scales)
```

##Factors

###Part 1
Factors, like other vectors, can be manipulated with `dplyr` functions like arrange(). However, applying the arrange() function to factors doesn't alter their position in a `ggplot` figure. If we wanted to visualize life expectancy in Asian countries in 2007, the figure looks the same regardless of whether the data frame is arranged by life expectancy or country name. These changes only apply to the tabulated data.


```{r}
nlevels(gapminder$country)

americas_1997 <- gapminder %>% 
  filter(continent == "Americas", year == 1997)



americas_1997 %>% 
  arrange(gdpPercap) %>% 
  ggplot(aes(gdpPercap, country)) +
  geom_point()

#or
americas_1997 %>% 
  arrange(country) %>% 
  ggplot(aes(gdpPercap, country)) +
  geom_point()
```

Using the `forcats` package, factors can be manipulated in some useful ways. We can reorder the above figure to better represent the data.

```{r}

americas_1997 %>% 
    mutate(country = fct_reorder(country, gdpPercap)) %>% 
    ggplot(aes(gdpPercap, country)) + 
    geom_point()

  
```

We can see that, even with the countries outside of the Americas filtered off, the levels corresponding to those countries are still represented in the factor.
```{r}
americas_1997$country %>% 
  nlevels()
americas_1997$country %>% 
  head()
```

We can drop the unused levels using fct_drop() to retain levels in the factor "country" and the factor "continent" that pertain to the Americas.

```{r}
americas_1997 %>% 
  mutate(country = fct_drop(country)) %>% 
  mutate(continent = fct_drop(continent)) %>% 
  str()
```



###Part 2 
For writing and re-reading a data set, we will use the gapminder data with its levels reordered and dropped as they were in Part 1.


```{r}
reordered_americas_1997 <- americas_1997 %>% 
  mutate(country = fct_drop(country)) %>% 
  mutate(continent = fct_drop(continent)) %>% 
  mutate(country = fct_reorder(country, gdpPercap))

```
Then we need to write reordered_americas_1997 to a csv

```{r}
write_csv(reordered_americas_1997, "americas_1997.csv")
```

and read it back into RStudio to see if reordered factors and dropped levels are still reordered and dropped.

```{r}
csv_americas_1997 <- read_csv("americas_1997.csv")
```

Comparing csv_americas_1997 and reordered_americas_1997:

```{r}
csv_americas_1997 %>% 
  str()
reordered_americas_1997 %>% 
  str()

identical(csv_americas_1997, reordered_americas_1997)
```
The 2 dataframes diverged when reordered_americas_1997 was written to a csv and read back into RStudio. The factors (country and continent) became character vectors and are no longer identical to the original vectors. Numeric vectors were unaffected in the process and remain identical.

```{r}
identical(csv_americas_1997$gdpPercap, reordered_americas_1997$gdpPercap)
```

###Part 3

####Visualization of figures

Looking at the figure from Part 1, it's evident that it could communicate more information with its visual presentation.
```{r}
americas_1997 %>% 
    mutate(country = fct_reorder(country, gdpPercap)) %>% 
    ggplot(aes(gdpPercap, country)) + 
    geom_point()
```

By adjusting certain elements, the figure presents more information to the reader about the dataset.

```{r}

(americas_plot <- americas_1997 %>% 
    mutate(country = fct_reorder(country, gdpPercap)) %>% 
    ggplot(aes(gdpPercap, country)) + 
    geom_point(aes(colour = lifeExp, size = pop)) +
    scale_x_continuous(labels = dollar_format()) +
    scale_color_distiller(palette = "Reds") +
    labs(x = "GDP Per Capita",
       y = "Country",
       title = "GDP per capita in the Americas"
  ) +
  theme_bw())


  
```

####Plotly
Using `plotly` allows a figure to have more interactivity and displayable data. By making the above figure a plotly graph, we can see the GDP per capita, population, and life expectancy of each data point just by hovering over it. If the graph was more crowded with data, plotly allows for zooming in on particular portions of the graph. 

```{r}
#ggplotly(americas_plot)     commenting for presentation on github
```

###Saving Figures
Figures can be saved as separate files using the `ggsave` function.
```{r}
ggsave("americas_1997.png", americas_plot)
```



